---
title: "WUP_04"
author: "Davy D"
date: "Spring 2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSS= read.csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We are trying to predict the WORDSUM based on schooling DEGREE.

WORDSUM is a numerical variable 
DEGREE is a categorical variable.

So we use the Numerical~Cat helper file. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

```{r}
boxplot(WORDSUM~DEGREE,data=GSS)
```

## Descriptive Results

The boxplot shows the people who scored the highest are Bachelor, Graduate, and Junior College and the ones that scored the lowest are HighSchool and LT High school. 

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


### Numerical Descriptive Results


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSS)
```

The table shows that, the Graduate and Bachelor have the highest average performance, and High school and LT highschool have the lowest average performance.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSS)
summary.aov(model1)
```

The P-value is 0.000236 which is less than 0.05 we reject the null hypothesis we can conclude there is some difference there is some differences in wordsum scores based on level on education. 

```{r}
plot(TukeyHSD(model1))
```


The biggest differences are between graduates and lt high school and bachelors and lt high school. 


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The question if there was a difference between word some and degree level earned, we have effectively bee able to prove that there is a difference between the the WORDSUM and level of DEGREE. As proven in the above graphs the Graduate and Bachelor level degrees have a higher level of wordsum compared to those in the education level of LT highschool or Highschool.. also with a p-value of 0.000236 which is less than 0.05 we where able to reject the null hypothesis proving there is a difference in wordsum and degree level.
